---
title: "Exploration of Epimutations"
author: "Carlos Ruiz"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Summary

We run our algorithm to detect epimutations. In this document, we will plot all the candidate regions to select those more likely to harbour an epimutation.

```{r, echo = FALSE}
library(minfi)
library(ggplot2)
library(dplyr)
library(tidyr)
library(cowplot)

vers.epiMut <- "v3"
vers.QC <- "v2"
load(paste0("../results/methylation/finalQC_files/", vers.QC, "/gset.autosomic.Rdata"))
bumps <- read.delim(paste0("../results/methylation/Epimutations/", vers.epiMut, "/bumpsAll.txt"), 
                    header = FALSE)
colnames(bumps) <- c("chromosome", "start", "end", "value", "area", "cluster", "indexStart",
                     "indexEnd", "L", "clusterL", "Sample", "Anova")
```


# Plots

```{r}
makePlot <- function(row, bumps){

  sample <- bumps[row, "Sample"]
  idx <- bumps[row, "indexStart"]:bumps[row, "indexEnd"]
  
  beta <- data.frame(getBeta(gset[idx, ]))
  beta$pos <- start(rowRanges(gset[idx, ]))
  beta %>%
    gather("Sample", "Methylation", seq_len(ncol(gset))) %>%
    mutate(Sample = ifelse(Sample == "X16A66", "16A66", Sample)) %>%
    ggplot(aes(x = pos, y = Methylation, color = Sample == sample, group = Sample)) +
    geom_point() + 
    geom_line() +
    ggtitle(paste(sample, bumps[row, "chromosome"], bumps[row, "start"], bumps[row, "end"])) +
    theme_bw() +
    scale_y_continuous(limits = c(0, 1)) +
    theme(legend.position = "none",
           plot.title = element_text(hjust = 0.5))
}
```

G2CS8CO has a duplication in chromosome 15. This region contains some imprinted regions. The following imprinted regions present a different methylation pattern in G2CS8CO (mean methylation around 75%) than in the rest of samples (mean methylation around 50%).
 
```{r}
impBumps <-  subset(bumps, Sample == "G2CS8CO" & chromosome == "chr15")
rows <-  c(4, 8, 2, 3) 
plots <- lapply(rows, makePlot, bumps = impBumps)
plot_grid(plotlist = plots, nrow = 2, ncol = 2)
```



