---
title: "RNAseq OUTRIDER QC"
author: "Carlos Ruiz"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    number_sections: true
    toc: yes
    fig_caption: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, echo = FALSE}
library(OUTRIDER)
library(ggplot2)
library(dplyr)

RNA_vers <- "v3/"
outrider_folder <- paste0("../results/RNAseq/OUTRIDER_aberrations/", RNA_vers)
load(paste0(outrider_folder, "OUTRIDER_QC.Rdata"))

RNA_folder <- paste0("../results/RNAseq/Bioc_objects/", RNA_vers)
load(paste0(RNA_folder, "RNAseq_RangedSE_autosomicGenes.Rdata"))
```

This report contains the quality control of OUTRIDER results.

# Gene filtering

Genes were filtered out based on OUTRIDER parameters. A total of `r nrow (rse) - nrow(ods.filt)` genes were removed so a total of `r nrow(ods.filt)` genes were included in the analysis.

```{r}
plotExpressedGenes(ods.filt)
```

# Control for Confounders

```{r}
plotCountCorHeatmap(ods.filt, colGroups = c("Status"),
                    normalized = FALSE)
```

Before normalization, samples are clustered by status.

```{r}
plotCountCorHeatmap(ods.cor, colGroups = c("Status"), normalized = TRUE)
```

Some correlations still persist but the samples are not clustered by status.