---
title: "Samples descriptives"
author: "Carlos Ruiz"
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    number_sections: true
    toc: yes
    fig_caption: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(ggplot2)
library(SummarizedExperiment)

pheno_vers <- "v3/"
rnaseq_vers <- "v3/"

## load files
load(paste0("../results/phenotypes/", pheno_vers, "pheno.Rdata"))

## Remove G3CS5CO
pheno <- subset(pheno, SampleID != "G3CS5CO")

final_folder <- paste0("../results/RNAseq/Bioc_objects/", rnaseq_vers)
load(paste0(final_folder, "RNAseq_RangedSE_autosomicGenes.Rdata"))

## Define function for summary statistics
getSum <- function(vec, type = "continuous"){
  if (type == "continuous"){
    c(median = median(vec, na.rm = TRUE), 
      range = quantile(vec, probs = c(0.25, 0.75), na.rm = TRUE))
  } else if (type == "categorical"){
    t <- table(vec)
    data.frame(Categories = names(t),  N = as.vector(t), 
               Percentage = sprintf("%.2f", as.vector(prop.table(t)*100)),
               stringsAsFactors = FALSE)
  }
}

cats <- c("RNAseq", "Whole dataset")
```

# RNAseq population

This section describes the general features of the samples included in the RNAseq analysis. We will compared it with the proportions in the general population.

## Sex

```{r}
rbind(c("RNAseq", " ", " "),
      getSum(rse$Sex, type = "categorical"),
      c("Whole dataset", " ", " "),
      getSum(pheno$Sex, type = "categorical"))

```

We have a balanced number of males and females, similarly to the whole dataset.

```{r}
chisq.test(rbind(table(rse$Sex), table(pheno$Sex)))
```

RNAseq and whole dataset do not have a different proportion of males and females. 

## Gestational age

Gestational age is measured in weeks.

```{r}
wilcox.test(rse$GestAge, pheno$GestAge)
```

The median gestational age of RNAseq samples is 21 weeks (IQR: 20-22), the same than for the whole dataset. Thus, we did not find differences in gestational age between RNAseq and whole dataset.

## Status

```{r}
rbind(c("RNAseq", " ", " "),
      getSum(rse$Status, type = "categorical"),
      c("Whole dataset", " ", " "),
      getSum(pheno$Status, type = "categorical"))
```

We included all our controls in the RNAseq analysis, so the proportion of controls raised to 22%. 

```{r}
fisher.test(rbind(table(rse$Status), table(pheno$Status)))
```

The differences in the proportion of controls is not statistically significant.

## CHD group

```{r}
rbind(c("RNAseq", " ", " "),
      getSum(rse$pathClass, type = "categorical"),
      c("Whole dataset", " ", " "),
      getSum(pheno$pathClass, type = "categorical"))
```

As in the whole dataset, groups G2 and G4 comprised most of the samples. Groups G1 and G5 are only represented with one sample each.

```{r}
fisher.test(rbind(table(rse$pathClass), table(pheno$pathClass)))
```

The proportions of samples in each pathologic group is similar to the whole dataset.

## Genetic Cause

```{r}
rbind(c("RNAseq", " ", " "),
      getSum(rse$GeneticCause, type = "categorical"),
      c("Whole dataset", " ", " "),
      getSum(pheno$GeneticCause, type = "categorical"))
```

The proportion of samples with identified genetic cause is similar for RNAseq samples than in the whole dataset.

```{r}
#fisher.test(rbind(table(rse$GeneticCause), table(pheno$GeneticCause)))
```

We did not find differences either for this variables between RNAseq and the whole dataset.

## Sample Batch

```{r}
rbind(c("RNAseq", " ", " "),
      getSum(rse$SampleBatch, type = "categorical"),
      c("Whole dataset", " ", " "),
      getSum(pheno$SampleBatch, type = "categorical"))
```

Almost all samples from RNAseq come from sample batch VHIR1. The other batches only contributed with one sample each. 