---
title: "ValidateEpisignatures"
author: "Carlos Ruiz"
date: "4/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Genetic disease might be produced by a general disregulation of DNA methylation. We explored in our dataset if the different CHD subgroups had different methylation patterns.

# Episignatures definition

We adapted the algorithm from PMID: 30929737 to define episignatures:

1. Divide dataset 3/4 train, 1/4 test. We sampled independently each CHD class
2. Run limma of groups each against the rest
3. Select top 1,000 CpGs with minimum difference > 0.1
4. Train SVM with CpGs selected in any comparison
5. Run SVM in test dataset
6. Repeat 4 times, to include all samples in test

As a result, all samples were train 3 times and test 1. 

```{r}
library(tidyr)
library(dplyr)
library(minfi)
library(ggplot2)
library(dendextend)


load("../results/methylation/Episignatures/cohort.training.Rdata")
load("../results/methylation/finalQC_files/v4/gset.autosomic.Rdata")
load("../results/methylation/Episignatures/GSE62629.results.Rdata")
load("../results/methylation/Episignatures/GSE62629.GenomicRatioSet.EPICcpgs.Rdata")


pred.list <- lapply(episig_subset, function(x) x$pred_test)

pred.df <- data.frame(predClass = unlist(pred.list),
					  set = rep(1:4, lengths(pred.list)),
					  sample = names(unlist(pred.list))) %>%
	mutate(trueClass = gset[, sample]$pathClass,
		   match = predClass == trueClass)

table(pred.df$predClass, pred.df$trueClass)
```

Individuals classified as having other malformations are correctly classified in all sets. Next, there is a great confussion between Conotruncal Malformations and Left heart hypoplasia. Therefore, we decided to consider both categories as equivalent and repeat the analysis:

```{r}
pred.list_merge <- lapply(episig_subset_mergeClass, function(x) x$pred_test)

pred.df_mrege <- data.frame(predClass = unlist(pred.list_merge),
					  set = rep(1:4, lengths(pred.list_merge)),
					  sample = names(unlist(pred.list_merge))) %>%
	mutate(trueClass = gset[, sample]$pathClass,
		   match = predClass == trueClass)

table(pred.df_mrege$predClass, pred.df_mrege$trueClass)
```

After merging both categories, 41/44 (93%) of samples were correctly classified, with the exception of 2 controls and 1 individuals with other malformations that were incorrectly labeled as Conotruncal Malformations/Left heart hypoplasia.

We repeated the whole process but using the whole cohort as training. 

```{r}
mat <- getBeta(gset[svm_all$feats, ])

cols <- c("#000000", "#CC6666", "#9999CC", "#66CC99")
names(cols) <- c("Control", "Conotruncal Malformations", 
				 "Left heart hypoplasia", "Other malformations")

colPlot <- cols[gset$pathClass]

heatmap(mat, Rowv = NA, scale = "none", 
		col = gray.colors(100, start = 0, end = 1),
		ColSideColors =colPlot)
```

The features included in the classifier, clearly separated controls (black) from cases. Then, individuals with other malformations are separated from Conotruncal Malformations and left heart hypoplasia.

# Validation in MBD-seq dataset

We tried to validate our SVM classifier in a GEO dataset. To this end, we selected the positions in the MBD-seq matching the CpGs in the EPIC array and run the SVM classifier:

```{r}
table(pred.GSE62629)
```

All individuals were classified as Conotruncal Malformations. 

To explore this effect, we run a PCA on the subset of CpGs included in the SVM and including the measurements with the array and the MBD-Seq:

```{r}
pc_comb$x %>%
	data.frame() %>%
	mutate(Technique = factor(rep(c("MBD-Seq", "EPIC Array"), c(ncol(gset.GSE62629.filt), ncol(gset))))) %>%
	ggplot(aes(x = PC1, y = PC2, col = Technique)) +
	geom_point() + 
	theme_bw()
```

Samples are clustered by the technique, so the SVM trained with the EPIC array cannot be directly applied to the MBD-Seq dataset.

Despite this, we aimed to test whether the features selected with our approach could be used to separate samples by class. 


```{r}
mat.GSE <- data.matrix(getBeta(gset.GSE62629.filt[svm_all$feats, ]))

cols2 <- c("#000000", "blue", "green")
names(cols2) <- c("normal", "Tetralogy of Fallot", 
				 "ventricular septal defect")

colPlot2 <- cols2[gset.GSE62629.filt$"disease state:ch1"]

heatmap(mat.GSE, Rowv = NA, scale = "none", 
		col = gray.colors(100, start = 0, end = 1),
		ColSideColors = colPlot2)
```

We see some clustering by pathologic class, with control samples (black) on one branch, Tetratology of Fallot (blue) in the middle and ventricular septal defect (green) in the middle.

Finally, we run a PCA to the CpGs in MBD-Seq dataset common in EPIC and to the features selected in the SVM classifier and compared the clustering.

```{r cars}
hcGSE <- as.dendrogram(hclust(dist(pc_GSE$x)))
hcGSE_svm <- as.dendrogram(hclust(dist(pc_GSE_svm$x)))

ord1 <- order.dendrogram(hcGSE)
labels_colors(hcGSE) <- cols2[gset.GSE62629.filt[, ord1]$"disease state:ch1"]
plot(hcGSE, main = "All features")

ord2 <- order.dendrogram(hcGSE_svm)
labels_colors(hcGSE_svm) <- cols2[gset.GSE62629.filt[, ord2]$"disease state:ch1"] 
plot(hcGSE_svm, main = "SVM features")
```

Samples are more clustered by pathologic state when restricting the PCA to the features in the SVM classifier.

# Features in the literature

```{r}
csvs <- dir(pattern = "SignificantCpGs", path = "../data", full = TRUE)
tabList <- lapply(csvs, read.csv, header = TRUE)
tabList <- lapply(tabList, function(x) {
	colnames(x) <- x[1, ]
	colnames(x)[1] <- "TargetID"
	x <- x[-1, ]
		}
	)
tabList[[9]] <- tabList[[9]][-1, ]

cpgList <- lapply(tabList, function(x) x$TargetID)

comCpGs <- sapply(cpgList, function(x) intersect(x, svm_all$feats))
comCpGs

feats.sub <- unlist(lapply(episig_subset, function(x) x$feats))
comCpGs_sub <- sapply(cpgList, function(x) intersect(x, feats.sub))
comCpGs_sub
```

We observed very few overlap between the probes in SVM and the probes previously reported in the literature. Nonetheless, we used a different array and a different tissue.
