---
title: "Methylation GATA4 vs inversions"
author: "Carlos Ruiz"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Objective

Test differences in GATA4 DNA methylation due to inv8p23.1 genotypes.

# Load data

```{r}
library(minfi)
library(limma)
library(dplyr)
library(scoreInvHap)
library(ggplot2)
library(cowplot)

load("../../results/methylation/finalQC_files/2021-03-08/gset.autosomic.Rdata")
gata4_cpgs <- read.table("../../data/Natalia_cpgs_gata4.txt")$V1
load("../../scripts/Natalia_ISGlobal/inv8Genos.scoreInvHap.Rdata")
```

# Add inversions to methylation
```{r}
inv8class <- classification(inv8sc)
gset$inv8 <- factor(inv8class[colnames(gset)], levels = c("NN", "NI", "II"))
gset$inv8add <- as.numeric(gset$inv8) - 1
table(gset$inv8)
table(gset$inv8, gset$Status)

```


# Methylation vs inv8

Explore association between the inversion and the methylation. Adjust for sex and CHD status.

```{r}
model <- model.matrix(~ inv8add + Sex + Status, colData(gset))
lmfit <- lmFit(getBeta(gset[, rownames(model)]), design = model)
lmFite <- eBayes(lmfit)
tab <- topTable(lmFite, n = Inf, coef = 2)
tab$P.adj2 <- p.adjust(tab$P.Value)

```
```{r}
tab[rownames(tab) %in% gata4_cpgs, ]
```

Only two CpGs significantly different (P.adj2 < 0.05)

```{r}
mapCpgs <- gata4_cpgs[gata4_cpgs %in% rownames(gset)]

a <- lapply(mapCpgs, function(cpg){
  df <- data.frame(methy = as.vector(getBeta(gset[cpg, ])), inv =  gset$inv8, row.names = NULL)
  df <- df[!is.na(df$inv), ]
  ggplot(df, aes(x = inv, y = methy)) + 
    geom_boxplot() +
    ggtitle(cpg)
})
plot_grid(plotlist = a[1:12], nrow = 3, ncol = 4)
plot_grid(plotlist = a[13:20], nrow = 3, ncol = 4)

```

# Methylation vs interaction inv8 + CHD status

```{r}
model2 <- model.matrix(~ inv8add*Status + Sex, colData(gset))
lmfit2 <- lmFit(getBeta(gset[, rownames(model2)]), design = model2)
lmFite2 <- eBayes(lmfit2)
tab2 <- topTable(lmFite2, n = Inf, coef = 5)
tab2$P.adj2 <- p.adjust(tab2$P.Value)
tab2[rownames(tab2) %in% gata4_cpgs, ]
```

Three CpGs show a significant interaction (P.adj2 < 0.05)

```{r}

b <- lapply(mapCpgs, function(cpg){
  df <- data.frame(methy = as.vector(getBeta(gset[cpg, ])), inv =  gset$inv8, 
                   status = gset$Status, row.names = NULL)
  df <- df[!is.na(df$inv), ]
  ggplot(df, aes(x = inv, y = methy)) + 
    geom_boxplot() +
    ggtitle(cpg) +
    facet_grid(~ status)
})
plot_grid(plotlist = b[1:12], nrow = 3, ncol = 4)
plot_grid(plotlist = b[13:20], nrow = 2, ncol = 4)
```

