---
title: "Methylation GATA4 vs CHD"
author: "Carlos Ruiz"
date: "31/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Objective

Test differences in GATA4 DNA methylation between controls and CHD (congenital heart defects) patients.

# Load data

```{r}
library(minfi)
library(limma)
library(dplyr)

load("../../results/methylation/finalQC_files/2021-03-08/gset.autosomic.Rdata")
gata4_cpgs <- read.table("../../data/Natalia_cpgs_gata4.txt")$V1
```

# Run linear model

```{r}
pheno <- colData(gset)
pheno$Status <- factor(pheno$Status , levels = c("Control", "Case"))
model <- model.matrix(~ Status + Sex, pheno)
lmfit <- lmFit(getBeta(gset), design = model)
lmFite <- eBayes(lmfit)
tab <- topTable(lmFite, n = Inf, coef = 2)
```

# Natalia's list 

```{r}
tab[rownames(tab) %in% gata4_cpgs, ]
```

Only two CpGs significantly different

```{r}
gset.gata4 <- gset[rownames(gset) %in% gata4_cpgs, ]

mat <- getBeta(gset.gata4)
cols <- c("#000000", "#CC6666", "#9999CC", "#66CC99")
names(cols) <- c("Control", "Conotruncal Malformations", 
                 "Left heart hypoplasia", "Other malformations")

colPlot <- cols[gset$pathClass]
heatmap(mat, scale = "none", 
        col = gray.colors(100, start = 0, end = 1),
        ColSideColors = colPlot, cexCol = 2)
```
Controls are in black and colors are different CHD subtypes. No clustering by patient status. 

```{r}
par(mfrow = c(2, 3))
invisible(lapply(rownames(mat), function(cpg){
  df <- data.frame(cpg = as.vector(getBeta(gset[cpg, ])), status =  gset$Status)
  boxplot(cpg ~ status, main = cpg, df)
}))
```
For different CpGs, cases have higher methylation.

```{r}
par(mfrow = c(1, 1))
hist(tab[rownames(mat) , ] $logFC)
```

We observe more CpGs where methylation is higher in cases (negative -log2FC) than in controls. 

# CpGs in Clara's paper ####

```{r}
lit_cpgs <- c("cg09626984", "cg20279283", "cg01546563", "cg19172575", "cg13434842")
tab[lit_cpgs , ] %>% arrange(P.Value)
```

Only one CpGs is significant.

```{r}
heatmap(getBeta(gset[lit_cpgs, ]), scale = "none", 
        col = gray.colors(100, start = 0, end = 1),
        ColSideColors = colPlot, cexCol = 2)
```

No clustering by CHD status is observed. 

```{r}
par(mfrow = c(2, 3))
invisible(lapply(lit_cpgs, function(cpg){
  df <- data.frame(cpg = as.vector(getBeta(gset[cpg, ])), status =  gset$Status)
  boxplot(cpg ~ status, main = cpg, df)
}))
```

For all the CpGs, the methylation in cases is higher than in controls.

```{r}
hist(tab[lit_cpgs , ]$logFC)
```

We observe more CpGs where methylation is higher in cases (negative -log2FC) than in controls. 

# GATA4 region ####

We selected all the CpGs in the array inside the GATA4 region.

```{r}
gset.reg <- subsetByOverlaps(gset, range(rowRanges(gset.gata4)))
tab[rownames(gset.reg) , ] %>% arrange(P.Value) %>% filter(P.Value < 0.05)
```

8 CpGs have a significant assocations.

```{r}
heatmap(getBeta(gset.reg), scale = "none", 
        col = gray.colors(100, start = 0, end = 1),
        ColSideColors = colPlot, cexCol = 2)
```

No clustering by CHD type is observed.


```{r}
hist(tab[rownames(gset.reg) , ]$logFC)
```
We observe more CpGs where methylation is higher in cases (negative -log2FC) than in controls. 

# Summary

It seems that controls have lower methylation than cases in the region. Nonetheless, these differences are not strong enough to stratify the samples.